General Relativistic Ray Tracing for Neutrinos

### 2016-10-12 paper.tex
Fix broken citations

\section{Introduction}
Neutrinos are one of the dominant energy transport phenomena at play in
neutron star mergers, heating, cooling, and pushing the disrupted nuclear
matter.
In addition, they change the composition of the matter via charged current
interactions.
Because neutrinos scatter over length scales both large and small with
respect to fluid scales, accurate models require a neutrino treatment that
respects the freedom of the neutrino distribution functions to vary
drastically from equilibrium.

This is a challenging task in the environment of a merger,
which generally lacks any spatial symmetries.
The state of the art today employs a truncated moment formalism
\citep{fouc2015-m1_bhns, fouc2016-m1_nsns},
evolving the zeroth- and first-angular moments of the distribution function
(the energy density and flux respectively) for the zeroth-energy moment
(the total energy).
Recently, \cite{fouc2016-m1_evolve_n} have expanded their code to also evolve
the number densities of neutrinos, giving a more accurate estimate of the
average neutrino energies.
But with only two angular moments and one energy moment, the distribution
function has very limited angular and spectral information.

\subsection{What physics problems do we want to address?}
But many interesting unsolved problems require an accurate model of the neutrino
spectra and angular distribution.
With a model of the neutrino emission from a merger we can
1) examine neutrino effects on the nucleosynthesis of the ejected material
\citep{robe2016-nsbh_sph_nucleo},
2) explore the rich flavor oscillation physics likely to occur here
\citep{malk2012-mnr_1, malk2014-mnr_2, malk2015-mnr_3, zhu2016-mnr_disk,
  vaan2016-uncovering_mnr},
3) improve closure relations used in truncated moment evolution schemes, and
4) look at jet formation due to neutrino annihilation
\citep{birk2007-nunubar}.

@@

\subsection{What is new/better about this ray tracing formulation?}
We build upon ray tracing methods already in the literature
\citep{hari2010-nunubar_gr, ?}
by including an elastic scattering approximation
\todo{and perhaps inelastic}
and removing any assumptions about the spacetime geometry.
To incorporate elastic scattering in our method, we employ the lower-order
estimates of the distribution function evolved in a moment scheme.

\subsection{Why does scattering matter?}
It is essential that we include scattering because although elastic scattering
doesn't change the emitted spectrum, it can significantly modify neutrino
distributions in angle, and dilute the emitted spectrum over a larger emitting
surface \citep{yonglin_collab}, especially for heavy-lepton neutrinos.
Any phenomena that involve neutrino-neutrino interactions (neutrino oscillations,
and neutrino-antineutrino annihilation) can depend sensitively on the angular
distribution.
And a diluted spectrum can have a different effect on the nucleosynthesis
occuring in the ejected and irradiated material.

@@

\subsection{Emission Spectra}
\label{ssec:spectra}
The merger models are described in \cite{fouc2015-m1_nsbh,fouc2016-m1_nsns}.

>>>>

\section{Introduction}
Neutrinos are one of the dominant energy transport phenomena at play in
neutron star mergers, heating, cooling, and pushing the disrupted nuclear
matter.
In addition, they change the composition of the matter via charged current
interactions.
Because neutrinos scatter over length scales both large and small with
respect to fluid scales, accurate models require a neutrino treatment that
respects the freedom of neutrino distribution functions to vary
drastically from thermodynamic equilibrium.

This is a challenging task in the environment of a merger,
which generally lacks any spatial symmetries.
The state of the art today employs either a
or a truncated moment formalism
\citep{fouc2015-m1_nsbh, fouc2016-m1_nsns},
evolving the zeroth- and first-angular moments of the distribution function
(the energy density and flux respectively) for the zeroth-energy moment
(the total energy).
Recently, \cite{fouc2016-m1_evolve_n} have expanded their code to also evolve
the number densities of neutrinos, giving a more accurate estimate of the
average neutrino energies.
But with only two angular moments and one energy moment, the distribution
function has very limited angular and spectral information.

\subsection{What physics problems do we want to address?}
But many interesting unsolved problems require an accurate model of the neutrino
spectra and angular distribution.
With a model of the neutrino emission from a merger we can
1) examine neutrino effects on the nucleosynthesis of the ejected material
\citep{robe2016-sph_nu_nucleo},
2) explore the rich flavor oscillation physics likely to occur here
\citep{malk2012-mnr_1, malk2015-mnr_2, malk2016-mnr_3, zhu2016-mnr_nsns_remnant,
  vaan2016-uncovering_mnr},
3) improve closure relations used in truncated moment evolution schemes, and
4) look at jet formation due to neutrino annihilation
\citep{birk2007-nunubar, hari2010-gr_nunubar_collapsar, ?}.

@@

\subsection{What is new/better about this ray tracing formulation?}
We build upon ray tracing methods already in the literature
\citep{hari2010-gr_nunubar_collapsar, youn2012-gr_radiative_transfer,
  kova2011-gr_ray_tracing, ?}
by including an elastic scattering approximation
\todo{and perhaps inelastic}
and removing any assumptions about the spacetime geometry.
To incorporate elastic scattering in our method, we employ the lower-order
estimates of the distribution function evolved in a moment scheme.

\subsection{Why does scattering matter?}
It is essential that we include scattering because although elastic scattering
doesn't change the emitted spectrum, it can significantly modify neutrino
distributions in angle, and dilute the emitted spectrum over a larger emitting
surface \citep{pere2016-asl}, especially for heavy-lepton neutrinos.
Any phenomena that involve neutrino-neutrino interactions (neutrino oscillations,
and neutrino-antineutrino annihilation) can depend sensitively on the angular
distribution.
And a diluted spectrum can have a different effect on the nucleosynthesis
occuring in the ejected and irradiated material.

@@

\subsection{Emission Spectra}
\label{ssec:spectra}
The merger models are described in \cite{fouc2015-m1_nsbh, fouc2016-m1_nsns}.
